Stack Safety is a Security Property

### 2021-04-10 tex/main.tex
major pass on intro

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  What is stack safety? The phrase has been used to describe the protections
  offered by a variety of mechanisms for guarding against
  attacks that corrupt the call stacks of running programs, including micro-policies
  implemented using metadata tagging mechanisms such as PIPE. But for specification
  these implementations largely appeal either to intuitive notions of the behavior
  of a stack, or to specific examples of bad behavior.
  %
  We propose a formal definition phrased in the technical terminology of language-based
  security---as a combination of an integrity property (a caller's stack frame is
  invariant over calls) and a confidentiality property (callees lack knowledge of the
  contexts of their calls). These combine with a straightforward well-bracketedness
  condition on the program's control flow that ensures that callers and callees remain
  meaningfully distinct.
  %
  While our formalism is informed by the PIPE implementation, the primary properties reflect
  general principles of stack safety. We define additional properties tailored for specific
  contexts: a stronger property amenable to efficient testing and a weaker one that is
  approximated by more efficient, lazy policies.

@@

%% Keywords
%% comma separated list
\keywords{Stack Safety, Micropolicies}  %% \keywords are mandatory in final camera-ready submission

@@

The call stack is a perennial target for attacks
based on memory-safety vulnerabilities, with potential consequences
ranging from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a menagerie of
software and hardware techniques for stack safety have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\leo{Do we want to hunt for more refs here?}
\bcp{Sure, but later!}
\fi
This last provides an array of micro-policies offering protection at various granularities,
which offer a trade-off between protection and efficiently.

But what {\em is} the stack safety that these techniques enforce? Precise formal
definitions are mostly lacking: rather, the notion is commonly defined in a
negative way, by enumerating the bad things that a given enforcement mechanism
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
Indeed, we are only aware of one quite recent attempt to frame a positive definition, by
\citet{Skorstengaard+19}, which defines stack safety as the conjunction of two properties:
{\em local state encapsulation} and {\em well-bracketed control flow}.
These properties are formalized by defining a capability-based machine
that ``satisfies [stack safety] by construction.''
%
While this work represents an important step in the right direction, we believe
there is still room to craft an operational characterization
of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
In other words, we seek formal criteria that can be applied
to a mechanism for protecting stack-based low-level programs to judge
whether or not it actually ``guarantees stack safety.'' We begin with a strong
local state encapsulation property that closely models the behavior of eager
micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
is too strong to characterize stack safety; reasonable enforcement mechanisms
exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
to support this technique and argue that the new property, defined relative to an
arbitrary notion of observation. This property is still strong enough to prevent
the bad behaviors we associate with unsafe stack behavior, but weak enough that
other reasonable enforcement mechanisms can enforce it.

While our criteria are designed with PIPE micro-policies in mind, they are
agnostic about many implementation details and should be applicable to other
systems with minor adjustment.
%
Our key contribution is an operational model of local state encapsulation
using the technical framework of language-based security
\apt{this idea deserves more prominence}, with eager and lazy variants,
together with a simple characterization of well-bracketed control flow.

@@

\paragraph*{The Use of Stack Safety}

How does the programmer rely on stack safety? The stack makes implicit promises about
the behavior of function calls. First, that when a call is finished, the caller can continue
from where it left off with its local state unchanged. And second, that the call's
behavior is determined only by a predictable interface that does not include the caller's
local state.

%The most natural example is the first
%in figure \ref{fig:examples1}, in which {\tt main} calls stores 42 to a local variable {\tt x},
%calls {\tt f}, and then prints {\tt x}. Here the programmer should be able to expect that
%{\tt main} will print 42, but in C such an assumption may be wrong, if {\tt f} should
%overwrite {\tt main}'s stack data.

Both expectations imply predictable control flow, termed {\em well-bracketed
control flow} (WBCF) by \citeauthor{Skorstengaard+19}. Calls always jump to valid entry
points and return to their call sites. Then the data portion of the first expectation is for
{\em integrity}: that a caller's local data is protected from modification by its callee.
This is vital for local reasoning about the behavior of the caller. The security implications
are significant: without integrity, a caller that branches on a local variable could be
manipulated by its callee into unexpected behavior.

Alongside integrity, we have {\em confidentiality}, in which a caller's data is protected
from leakage by its callee. As a reasoning principle, confidentiality tells us that to
predict the behavior of a call we need only consider a subset of the machine state.
And in security terms, we can guarantee the secrecy of local data that we haven't passed
even when calling a badly behaved function. Together, integrity and confidentiality
comprise \citeauthor{Skorstengaard+19}'s {\em local state encapsulation} (LSE) property.

It must be reiterated that LSE is of little use without WBCF: without WBCF, a callee
could subvert LSE by ``returning'' to arbitrary code; after the return LSE should permit
accessing the caller's state, but the execution would remain under the callee's control.

To phrase LSE and WBCF as security properties requires an explicit attacker model:
a clear understanding of who is being protected and from what. Most existing
stack protection mechanisms envisage scenarios where a callee is
influenced by external inputs to corrupt the stack in some way. To
avoid getting into the details of exactly how this happens when
formalizing our properties, we use an even stronger model: when we
state the protections that a function can rely on when it makes a
call, we assume that the callee can execute arbitrary instructions
until it returns. Thus we have no need to differentiate safe code
from adversarial - all functions, and indeed all activations of the same function,
are mutually distrusting.
%
For this model to make sense, we need to know when a call or a return is taking
place. We assume that calls and returns are annotated in the code, so that
we can track the point at which control transfers from the protected caller
to the attacking callee -- and when that callee in turn becomes a protected
caller in its own right.

This ``nesting'' of properties is a key challenge in adapting the concepts
of language-based security to the stack setting. We address this issue by
turning any set of annotations into a dynamic model of {\em domains},
in which some stack locations are sealed by the caller that needs to preserve
them while the rest are unsealed and available for use by the active, attacking
callee.

First we introduce this model for a single simple stack with no sharing between callers and
callee. Then we expand it to allow passing of stack data, granting a callee one time
access to the caller's data. We do not model more sophisticated sharing, because to do
so realistically would require a separate protection scheme, such as a capability
model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
a separate policy outside the scope of stack safety. (A tag-based approach can easily
compose multiple such policies.)

Finally, we extend the model to a coroutine system featuring finitely-bounded stacks.
Though simplistic, this extension is a step toward concurrency and demonstrates that
the model is flexible enough to handle more sophisticated styles of control-flow.

After formally defining stack safety, we show that the definition is
enforceable using an existing tag-based enforcement
mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider
\citeauthor{DBLP:conf/sp/RoesslerD18}'s conservative Depth Isolation policy.
This policy testably enforces our {\em eager} properties, in which integrity is respected
at every step of a program, and each call respects integrity at each of its internal steps.
These properties are very strong, and our policies enforce them; this also makes them very
intensional, in that they do not distinguish internal behavior from external.
They also have the advantage that they are suitable for efficient testing, as they
fail immediately when a violation occurs.

The downside of the Depth Isolation policy is that it is rather expensive to implement.
In the same work, \citeauthor{DBLP:conf/sp/RoesslerD18} propose a different mechanism
that is more efficient by virtue of being lazy in its enforcement.
Instead of enforcing the integrity and confidentiality of the stack at the
exact moment when a callee returns, this policy signals violations when the
caller actually accesses data that the callee has written into its stack
frame. \ifaftersubmission\bcp{We could also hint at the lazy confidentiality
  property in the same way...}\fi
The lazy policy does not enforce our eager integrity property, and to our
knowledge there is no formal statement of the properties that it ought to enforce.
We propose {\em lazy} stack-safety properties, which capture the desired behavior of a
policy that detects integrity violations between when they occur and when they make
a visible impact. Intuitively this can be phrased as ``no visible reads after dangerous
writes.'' Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s existing
``Lazy Tagging and Clearing'' policy does not quite enforce this property, as we show in
testing, but formalizing the ideal property allows us to see how it falls short.
We provide a variant of ``Lazy Tagging and Clearing'' that does enforce lazy integrity
at some performance cost.
%We prove (in Coq) that the eager property implies the lazy one.\apt{do we?
%  and is it so interesting anyhow?}
We argue that any system purporting to enforce stack safety ought to obey at least lazy
stack safety, making it a plausible extensional characterization.

%\bcp{I worry that we are being a little
%  dishonest here: the stated advantage of the lazy policy is that it is more
%efficient, but our variant is {\em not} efficient!}

In summary, we offer the following contributions:
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?}
  A novel {\em stack-safety property} capturing the intuitive concepts of strictly enforced
  local state encapsulation and well-bracketed control flow~(\cref{sec:lse-and-wbcf}).
  % These properties rule out common examples of stack memory abuse.
\item
  A property-based randomized testing framework that demonstrates that this
  property can be enforced by the Depth Isolation policy of
  \citet{DBLP:conf/sp/RoesslerD18}, which is based on hardware
  tagging~(\cref{sec:conservative}).
\item
  A {\em lazy} variant of the stack-safety property that can be enforced if the
  Lazy Tagging and Clearing policy of \citet{DBLP:conf/sp/RoesslerD18} is enhanced
  with unique tags on each activation, plus a proof (in Coq) that this
  variant is weaker than the original~(\cref{sec:lazy}).
\item
  Standardized trace-based formulations of both eager and lazy confidentiality and integrity,
  parameterizeable by arbitrary control-flow models and assignments of machine components
  to {\em LSE domains}.
\item
  An LSE-domain model enhanced with passing of function arguments on the stack, with which our
  standardized trace properties extend stack safety with local sharing.
\item
  An LSE-domain model that applies to simple coroutining, with corresponding LSE
  properties and a {\em well-structured control flow} property that subsumes WBCF for each
  stack as well as guaranteeing legitimate coroutine control transfers.
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

  %\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
%  realizible by
%  a previously proposed micropolicy enforcement
%  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
%    substantial body of code}. We also present a variation of the
%  property + policy that is lazy, easier/more efficient to implement,
%  and provably weaker (\Cref{sec:lazy}).
%\item A provably stronger (inductive? step-by-step/stepwise) variant
%  of this property tuned for testing/verification.  We test the whole
%  thing with a cool framework. \bcp{Which section?}
\end{itemize}
\Cref{sec:relwork} discusses related work;
\cref{sec:future} sketches directions for future work.

We begin with a description of our threat model
(\cref{sec:threat}), a running example that we refer to throughout the paper
(\cref{sec:running-example}),  and some technical preliminaries
(\cref{sec:prelim}).

@@

This enforcement model directly describes micropolicies~\citep{pump_oakland2015}.
We believe that it is flexible enough to capture code-altering approaches, such as code
rewriting or capability-based techniques, perhaps with less elegance.

@@

\paragraph*{Micropolicies}
%
Micropolicies are a flexible tag-based, hardware-accelerated
reference monitoring mechanism; they been applied to stack safety
enforcement by \citet{DBLP:conf/sp/RoesslerD18}.
Here the policy state consists of metadata tags (e.g., identifying stack frames)
attached to each value in memory or registers, and the policy step
function checks that each machine operation obeys a set of rules on tags
(e.g., that the current $\PCname$ tag matches the tag on the stack
location being accessed), and halts the machine if not.  Distinguishing
ordinary machine state and policy state is very natural in this setting.
%% Here the machine state is extended with tags paired with all components of the
%% system, and a monitor checks for each operation that the tags on the operands
%% obey a set of rules. One could treat tags directly as part of the machine
%% state, and build the rules into the step function. But local state
%% encapsulation properties as we frame them are not intuitive in their
%% interaction with tags or other additional enforcement state, so it is helpful
%% to separate tags for clarity. This also distinguishes a hardware fail-stop from
%% a software fail state.
%% \rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
%%   policy? }

@@

Up til now we have assumed that function arguments and returns are passed solely in registers.
This is an oversimplication, which we will now correct. There are important stack safety concerns
when passing arguments on the stack, as seen in example \ref{fig:passing}. In this example,
{\tt main} passes a reference to {\tt x}, but then clears {\tt x} and instead passes one to
{\tt y}. Before each call, when the address of a variable is taken, we add a new annotation
\(P(x)\) or \(P(y)\) identifying the memory to be passed, and our stack model marks it in
magenta. \apt{Why are there two line numbers assoicated with the two calls?}
Then at the call, the parts of the caller's frame that are not passed are sealed, as usual. Passed
memory also tracks which depth owns it, and it may only be accessed by the immediate callee
(at its depth plus one).

@@

%\paragraph*{Micropolicies}
%
The enforcement policy of \citeauthor{DBLP:conf/sp/RoesslerD18} relies
on a programmable, tag-based reference monitor that runs alongside the
program.  To control this monitor, they use a programming model that
allows fine-grained manipulation of metadata tags to encode so-called
\emph{micropolicies}~\citep{pump_oakland2015}.
In such a system, all values in
memory addresses and registers (including the $\PCname$)
are enriched with an abstract metadata tag, which can represent
arbitrary information about the value. A micropolicy is defined as a set of tags and a
collection of software-defined rules, indexed by machine opcode.
At each step of the machine, the relevant rule is applied to the tags on
the instruction's inputs (PC, registers, memory) and on the instruction itself, and produces
one of two outcomes: either the instruction is allowed to execute (and generates
tags for the result of the operation and the new PC), or the machine
fail-stops with a policy violation.
\citeauthor{pump_oakland2015} have shown that a wide range of micropolicies can
be defined using this scheme.

Efficient execution of these micropolicies relies on hardware implementations,
such as the PUMP architecture~\citep{pump:asplos2015}.  Tags are represented
as word-size bit vectors, stored separately from the regular memory and registers.
The hardware incorporates a rule cache
to allow quick retrieval of rule outputs for mapped inputs. If the cache misses,
the hardware traps to a software handler (running in a privileged context or
on a co-processor) to compute the rule result. To obtain adequate performance,
it is important to design micropolicies so that they hit in the cache as
much as possible. Thus, practical policies maintain a small working set of
distinct tags.
%% %
%% \rb{Maybe go into more detail, examples later\ldots or try to segue into stack
%% policies}

\subsection{A Conservative Tag Policy}
\label{sec:conservative}
%
The first micropolicy we present to enforce the definition of stack
safety developed in \cref{sec:lse-and-wbcf} is a small variation
on the Depth Isolation policy presented by
\citet{DBLP:conf/sp/RoesslerD18}. Their policy tags the stack memory
with ownership information associated to each stack frame (and to each
separate object inside that frame, a more finely grained access
control that we do not need to consider here), and tags registers containing
stack pointers with access permission information. The policy also
uses tags on instructions to identify the code
sequences that have permission to manage the stack, say during calls and returns.
%% \rb{Conceivably on alloc
%%   operations, etc., which we have not yet mentioned.}

Our micropolicy works as follows. (We assume for ease of exposition that no
arguments are passed on the stack.)
%% \rb{Come back to this later?}.
The micropolicy maintains tags on values in memory of the form $\tagStackDepth{n}$,
indicating locations that belong to the stack frame at activation depth $n$,
or $\tagNoDepth$; it tags the $\PCname$ with tags of the form $\tagPCDepth{n}$.
During normal execution, the micropolicy rules only permit load and store operations
when the target memory is tagged with the same depth as the current {\PCname} tag.
Initially, the entire stack is tagged $\tagNoDepth$,
%\rb{actually, the ``initial frame'' should be tagged with its proper depth}
and the {\PCname} has tag $\tagPCDepth{0}$. These tags are altered
at exactly those points in
the program where the contour of the stack-safety property changes:

@@

This discipline suffices to enforce local stack encapsulation.
To implement it,  we define ``blessed'' instruction sequences
intended to appear at the entry and exit of each function,
which manipulate tags as just described in addition to performing the
usual calling convention tasks of saving/restoring the return address to/from
the stack and adjusting the stack pointer. With the aid of an additional tag on
the $\rsp$ register, these sequences also serve to enforce
well-bracketed control flow.
The micropolicy guarantees atomic execution of these sequences
using a combination of tags on the instructions
and an additional tag on the $\PCname$; we omit the details here.

There remains the question of how to ensure that the sequences are
invoked at the right places. For a program to enjoy the stack-safety property
(with respect to a particular call map), all we need require is that
the entry sequence be initiated at any instruction marked as a call
in the call map. This is achieved by giving these instructions (another) special tag,
and it is easy to check statically that this has been done correctly.
If the code fails to initiate an exit
sequence at a point where the stack-safety property expects a return,
the micropolicy will incorrectly behave as if execution
is continuing in the callee, but since the callee never has stronger access
rights than the caller, this is harmless.

@@

%% Only programs whose sequences are well-formed are
%% protected by the micropolicy, although programs are still protected
%% (by fail-stopping) even if the intended sequences are missing. By
%% tracking the execution of the blessed sequences in the tag state of
%% the {\PCname} register, the micropolicy ensures that the entry
%% sequences of calls declared in the call map are executed correctly.

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
\apt{maybe: ``Stack safety is a testable security property'' ?}a
  What is stack safety? The phrase has been used to describe the protections
  offered by a variety of mechanisms for guarding against
  attacks that corrupt the call stacks of running programs, including micro-policies
  implemented using metadata tagging mechanisms such as PIPE. But for specification
  these implementations largely appeal either to intuitive notions of the behavior
  of a stack, or to specific examples of bad behavior.
  %
  We propose a formal definition phrased in the technical terminology of language-based
  security---as a combination of an integrity property (a caller's stack frame is
  invariant over calls) and a confidentiality property (callees lack knowledge of the
  contexts of their calls). These combine with a straightforward well-bracketedness
  condition on the program's control flow that ensures that callers and callees remain
  meaningfully distinct.
  %
  While our formalism is informed by the PIPE implementation, the primary properties reflect
  general principles of stack safety. We define additional properties tailored for specific
  contexts: a stronger property amenable to efficient testing and a weaker one that is
  approximated by more efficient, lazy policies.

@@

%% Keywords
%% comma separated list
\keywords{Stack Safety, Micro-Policies}  %% \keywords are mandatory in final camera-ready submission

@@

The call stack is a perennial target for attacks
based on memory-safety vulnerabilities, with potential consequences
ranging from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a menagerie of
software and hardware techniques for stack safety have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\leo{Do we want to hunt for more refs here?}
\bcp{Sure, but later!}
\fi
% 
The implementation and configuration of these mechanisms can be complex, 
and often involves making compromises that trade off protection for efficiency.  
Thus it would be very useful to have precise formal specifications for stack safety
that give us a basis for comparing the security guarantees claimed by different enforcement
techniques, and for verifying that these claims are met by particular implementations.

For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
In essence, memory locations within each stack frame are tagged with the identity of
the function that owns the frame, and access to those locations is then gated
according to the function that is currently executing. Many alternatives can
be implemented on top of this basic idea, by choosing how to identify functions
and deciding just which accesses are checked.  These alternatives vary in their
efficiency, but also in the spatial and temporal precision of their security guarantees.
For example, the \emph{Depth Isolation} policy strictly requires that a function owns any
frame into which it writes; this seems simple and intuitively correct, but can be expensive to implement.
On the other hand, the \emph{Lazy Tagging and Clearing} policy improves performance by 
allowing illicit writes to the frame so long as the written locations are not
subsequently read. The intuition justifying laziness is that any genuinely dangerous writes will ultimately be
detected before they can affect program behavior.  But we clearly need a more formal foundation for
expressing, comparing and verifying these micro-policy security claims, as well as those of
other enforcement technologies.

So how should we formally define stack safety? Definitions in the literature are commonly informal
and proceed by enumerating the specific bad things that a given enforcement mechanism
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
The most abstract characterization of which we are aware is in 
recent work by \citet{Skorstengaard+19}, which defines stack safety as the conjunction of two properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read and written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the point just after their call (or don't return at all). 
\citet{Skorstengaard+19} formalize these properties by defining an idealized machine
in which they visibly hold by construction; an enforcement mechanism (in their case, hardware capabilities)
can then be shown to provide stack safety by giving a fully abstract embedding function
from the idealized machine to an ordinary machine equipped with the enforcement.

In this work, we take the (informal) definitions of WBCF and LSE as a starting point,
but choose a different approach to formalizing them, based on the
technical framework of language-based security~\citep{??}.  We define
trace-based properties of \emph{stepwise stack integrity} and \emph{stack confidentiality}.
In applying these security concepts, we treat each caller-callee pair as mutually distrustful
and view the caller as needing protection from the callee. \apt{Need more explanation of
  ``nesting'' and our novelty factor here? Also about domains??}
{\em Integrity} says that a caller's local data is protected from modification by its callee.
This is vital for local reasoning about the behavior of the caller. 
{\em Confidentiality} states that a caller's data is protected
from being leaked into its callee, and also that the callee cannot see uninitialized data.
As a reasoning principle, confidentiality guarantees that the callee's
behavior is determined only by a predictable interface that does not include the caller's
local state or the uninitialized portion of the stack.
Together, these properties (informally) capture LSE, as well as the ``temporal stack safety''
notion that data outside of live frames should be inaccessible (which LSE does not address).
We also give a simple trace-based formalization of WBCF, which is essential for keeping the
notions of ``caller' and ``callee'' sensible.

To illustrate the applicability of these properties, we perform randomized testing to
validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces them.
\apt{more to say?}

However, their Lazy Tagging and Clearing micro-policy does not enforce stepwise integrity.
To address this, we develop an \emph{observational stack integrity} property that 
is intended to specify enforcement mechanisms that 
detect integrity violations between the point when they occur and the point when they make
an observable impact. Intuitively this property can be phrased as ``no visible reads after dangerous
writes.''  We prove (in Coq) that stepwise integrity implies observational integrity.

Unfortunately the existing Lazy Tagging and Clearing 
micro-policy does not quite enforce this property either, as we show in
testing~\apt{really??}, but formalizing the ideal property allows us to see how it falls short.
We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce observational integrity,
at some performance cost.\apt{but we're not testing that, right?}

Observational integrity is attractive in being significantly more \emph{extensional} than
stepwise integrity.  We argue that it may therefore be a better minimal specification for
stack safety enforcement mechanisms in general.

Our initial description of the properties assumes single simple stack with no sharing between callers and
callees.  We expand it to allow passing of scalar stack data, granting a callee 
access to the caller's data. We do not model more sophisticated sharing, because to do
so realistically would require a separate protection scheme, such as a capability
model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
a separate policy outside the scope of stack safety. (A tag-based approach can easily
compose multiple such policies.)

Finally, we define integrity and confidentiality properties for a coroutine system featuring finitely-bounded stacks.
Though simplistic, this extension is a step toward concurrency and demonstrates that
our approach is flexible enough to handle more sophisticated styles of control-flow.

While our property definitions were developed with micro-policy enforcement in mind, they
agnostic about many implementation details and should be applicable to other
systems with minor adjustment.

@@

In summary, we offer the following contributions:
\begin{itemize}
\item 
  We harness the language-based security concepts of integrity and confidentiality 
  to formalize a parameterized {\em stepwise stack-safety property}, and argue that
  it captures the intuitive meaning of local stack encapsulation, well-bracketed control
  flow, and temporal stack safety~(\cref{sec:lse-and-wbcf}).
\item
  We apply a property-based randomized testing framework to provide evidence that 
  this stepwise property is enforced by the Depth Isolation policy of
  \citet{DBLP:conf/sp/RoesslerD18} (\cref{sec:conservative}).
\item 
  We formalize a weaker but more extensional \emph{observational} variant of the
  stack safety property, and argue that it accurately characterizes 
  the Lazy Tagging and Clearing policy of \citet{DBLP:conf/sp/RoesslerD18},
  when enhanced with unique tags on each activation~(\cref{sec:lazy}).
\item 
  We show how our models can be instantiated and extended to handle passing
  arguments on the stack~(\cref{sec:??}) and to support simple coroutining~(\cref{sec:??}).
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

  %\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
%  realizible by
%  a previously proposed micro-policy enforcement
%  mechanism~(\Cref{sec:micro-policy}). \leo{And it says yes on a
%    substantial body of code}. We also present a variation of the
%  property + policy that is lazy, easier/more efficient to implement,
%  and provably weaker (\Cref{sec:lazy}).
%\item A provably stronger (inductive? step-by-step/stepwise) variant
%  of this property tuned for testing/verification.  We test the whole
%  thing with a cool framework. \bcp{Which section?}
\end{itemize}
\Cref{sec:relwork} discusses related work;
\cref{sec:future} sketches directions for future work.

We begin with a description of our threat model
(\cref{sec:threat}), a running example that we refer to throughout the paper
(\cref{sec:running-example}),  and some technical preliminaries
(\cref{sec:prelim}).\apt{this para needs updating}

@@

This enforcement model directly describes micro-policies~\citep{pump_oakland2015}.
We believe that it is flexible enough to capture code-altering approaches, such as code
rewriting or capability-based techniques, perhaps with less elegance.

@@

\paragraph*{Micro-Policies}
%
Micro-Policies are a flexible tag-based, hardware-accelerated
reference monitoring mechanism; they been applied to stack safety
enforcement by \citet{DBLP:conf/sp/RoesslerD18}.
Here the policy state consists of metadata tags (e.g., identifying stack frames)
attached to each value in memory or registers, and the policy step
function checks that each machine operation obeys a set of rules on tags
(e.g., that the current $\PCname$ tag matches the tag on the stack
location being accessed), and halts the machine if not.  Distinguishing
ordinary machine state and policy state is very natural in this setting.
%% Here the machine state is extended with tags paired with all components of the
%% system, and a monitor checks for each operation that the tags on the operands
%% obey a set of rules. One could treat tags directly as part of the machine
%% state, and build the rules into the step function. But local state
%% encapsulation properties as we frame them are not intuitive in their
%% interaction with tags or other additional enforcement state, so it is helpful
%% to separate tags for clarity. This also distinguishes a hardware fail-stop from
%% a software fail state.
%% \rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
%%   policy? }

@@

Until now we have assumed that function arguments and returns are passed solely in registers.
This is an oversimplication, which we will now correct. There are important stack safety concerns
when passing arguments on the stack, as seen in example \ref{fig:passing}. In this example,
{\tt main} passes a reference to {\tt x}, but then clears {\tt x} and instead passes one to
{\tt y}. Before each call, when the address of a variable is taken, we add a new annotation
\(P(x)\) or \(P(y)\) identifying the memory to be passed, and our stack model marks it in
magenta. \apt{Why are there two line numbers assoicated with the two calls?}
Then at the call, the parts of the caller's frame that are not passed are sealed, as usual. Passed
memory also tracks which depth owns it, and it may only be accessed by the immediate callee
(at its depth plus one).

@@

%\paragraph*{Micro-Policies}
%
The enforcement policy of \citeauthor{DBLP:conf/sp/RoesslerD18} relies
on a programmable, tag-based reference monitor that runs alongside the
program.  To control this monitor, they use a programming model that
allows fine-grained manipulation of metadata tags to encode so-called
\emph{micro-policies}~\citep{pump_oakland2015}.
In such a system, all values in
memory addresses and registers (including the $\PCname$)
are enriched with an abstract metadata tag, which can represent
arbitrary information about the value. A micro-policy is defined as a set of tags and a
collection of software-defined rules, indexed by machine opcode.
At each step of the machine, the relevant rule is applied to the tags on
the instruction's inputs (PC, registers, memory) and on the instruction itself, and produces
one of two outcomes: either the instruction is allowed to execute (and generates
tags for the result of the operation and the new PC), or the machine
fail-stops with a policy violation.
\citeauthor{pump_oakland2015} have shown that a wide range of micro-policies can
be defined using this scheme.

Efficient execution of these micro-policies relies on hardware implementations,
such as the PUMP architecture~\citep{pump:asplos2015}.  Tags are represented
as word-size bit vectors, stored separately from the regular memory and registers.
The hardware incorporates a rule cache
to allow quick retrieval of rule outputs for mapped inputs. If the cache misses,
the hardware traps to a software handler (running in a privileged context or
on a co-processor) to compute the rule result. To obtain adequate performance,
it is important to design micro-policies so that they hit in the cache as
much as possible. Thus, practical policies maintain a small working set of
distinct tags.
%% %
%% \rb{Maybe go into more detail, examples later\ldots or try to segue into stack
%% policies}

\subsection{A Conservative Tag Policy}
\label{sec:conservative}
%
The first micro-policy we present to enforce the definition of stack
safety developed in \cref{sec:lse-and-wbcf} is a small variation
on the Depth Isolation policy presented by
\citet{DBLP:conf/sp/RoesslerD18}. Their policy tags the stack memory
with ownership information associated to each stack frame (and to each
separate object inside that frame, a more finely grained access
control that we do not need to consider here), and tags registers containing
stack pointers with access permission information. The policy also
uses tags on instructions to identify the code
sequences that have permission to manage the stack, say during calls and returns.
%% \rb{Conceivably on alloc
%%   operations, etc., which we have not yet mentioned.}

Our micro-policy works as follows. (We assume for ease of exposition that no
arguments are passed on the stack.)
%% \rb{Come back to this later?}.
The micro-policy maintains tags on values in memory of the form $\tagStackDepth{n}$,
indicating locations that belong to the stack frame at activation depth $n$,
or $\tagNoDepth$; it tags the $\PCname$ with tags of the form $\tagPCDepth{n}$.
During normal execution, the micro-policy rules only permit load and store operations
when the target memory is tagged with the same depth as the current {\PCname} tag.
Initially, the entire stack is tagged $\tagNoDepth$,
%\rb{actually, the ``initial frame'' should be tagged with its proper depth}
and the {\PCname} has tag $\tagPCDepth{0}$. These tags are altered
at exactly those points in
the program where the contour of the stack-safety property changes:

@@

This discipline suffices to enforce local stack encapsulation.
To implement it,  we define ``blessed'' instruction sequences
intended to appear at the entry and exit of each function,
which manipulate tags as just described in addition to performing the
usual calling convention tasks of saving/restoring the return address to/from
the stack and adjusting the stack pointer. With the aid of an additional tag on
the $\rsp$ register, these sequences also serve to enforce
well-bracketed control flow.
The micro-policy guarantees atomic execution of these sequences
using a combination of tags on the instructions
and an additional tag on the $\PCname$; we omit the details here.

There remains the question of how to ensure that the sequences are
invoked at the right places. For a program to enjoy the stack-safety property
(with respect to a particular call map), all we need require is that
the entry sequence be initiated at any instruction marked as a call
in the call map. This is achieved by giving these instructions (another) special tag,
and it is easy to check statically that this has been done correctly.
If the code fails to initiate an exit
sequence at a point where the stack-safety property expects a return,
the micro-policy will incorrectly behave as if execution
is continuing in the callee, but since the callee never has stronger access
rights than the caller, this is harmless.

@@

%% Only programs whose sequences are well-formed are
%% protected by the micro-policy, although programs are still protected
%% (by fail-stopping) even if the intended sequences are missing. By
%% tracking the execution of the blessed sequences in the tag state of
%% the {\PCname} register, the micro-policy ensures that the entry
%% sequences of calls declared in the call map are executed correctly.
